We have uploaded artifacts for evaluation, which contains all our code for running the experiments, survey questions, and the regex corpus used for analysis. 

Reviewers: thank you for your thorough and thoughtful feedback. Some doubts were raised about the survey, in particular, the size/variation of the sample and the precision of number of regexes composed per year.  We?d like to discuss these points briefly. 

- Regarding the sample size and variation, while the size is small (18 developers), all participants work in industry and have an average of 9 years of experience, so their opinions and experiences are likely to be diverse. Even so, replication with a larger and more diverse population would be very interesting, and in particular, we?d like to revamp the survey and include more qualitative questions to dive deeper into the common usage scenarios and frustrations with regexes. We?d also like to expand our artifact analysis to corroborate some of the survey findings by, for example, identifying regexes in any SQL queries embedded in source code. This work is a first step toward understanding regex usage, but there is so much left to do!  

- Regarding precision, we understand that these numbers are based on a brief recollection effort and should be treated accordingly, but even so, we were interested in gaining an understanding of how often developers use (or think they use) regexes. While precision remains a concern, these numbers do show that regexes are used quite frequently for a variety of purposes, and that is useful. It?s hard to think of a way to get these numbers more reliably, short of monitoring computer usage, but adding qualitative questions to the survey, as Reviewer 1 suggested, would be a good step in this direction. 
